Untitled

### 2016-07-27 inst/examples/flexdashboard.Rmd
ok

---
title: "A Quantitative Study of Finnish Knowledge Production `r params$min.year`-`r params$max.year`"
subtitle: "Subtitle here"
author: "Mikko Tolonen, Jani Marjanen, Hege Roivainen, Leo Lahti"
date: "`r Sys.Date()`"
fontsize: 13pt
geometry: margin=1in
linkcolor: magenta
urlcolor: blue
citecolor: black
institute: "University of Helsinki"
output: 
  beamer_presentation:
    theme: "boxes"
    colortheme: "orchid"
    fonttheme: "professionalfonts"
  pdf_document:
    fig_width: 7
    fig_height: 6
    fig_caption: true
    includes:
      in_header: header.tex
params:
  min.year: 
    label: "Start year:"
    input: numeric
    value: 1488
  max.year: 
    label: "End year:"
    input: numeric
    value: 1828
  journals:
    label: "Include journals:"
    value: TRUE
    input: checkbox
  catalogue:
    label: "Library catalog:"
    value: all
    input: select
    multiple: TRUE
    choices: [fennica, kungliga, all]
  language:
    label: "Language:"
    value: all
    input: radio
    choices: [finnish, swedish, all]
  data: 
    label: "Input dataset:"
    value: df.combined.Rds
    input: file
  slideparameter: 
    label: "Sliding value:"
    value: 100
    input: slider
    min: 0
    max: 1000
  date: 
    label: "Date:"
    value: !r date()
    input: date
  text: 
    label: "Text:"
    value: Some test text
    input: text
---

```{r 201606krakow-init, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE}

# For all parameter options for rmarkdown slide rendering, see:
# http://rmarkdown.rstudio.com/developer_parameterized_reports.html
# More header options in Pandoc manual
# http://pandoc.org/MANUAL.html#templates

# Read parameters from the header
mydate <- params$start
min.year <- params$min.year
max.year <- params$max.year

library(ggmap)
library(gisfin)
library(stringr)
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(bibliographica)
library(fennica)
library(sorvi)
library(devtools)
# install_github("ropengov/gisfin")
# devtools::install_github('cttobin/ggthemr')
#library(ggthemr)

knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(fig.path = "figure_slides/", dev="CairoPNG")
knitr::opts_chunk$set(fig.path = "figure_slides/")

@@

# Nice default themes
# https://github.com/cttobin/ggthemr#palettes
#ggthemr('fresh', text_size = 20)
# ggthemr('greyscale', text_size = 20)
#ggthemr('light', text_size = 20)
# ggthemr('pale', text_size = 20)
#ggthemr_reset() # Reset theme
```

```{r 201606krakow-init2, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE}
# Full combined catalogue (Fen + Kun) with marked duplicates
#df.combined.preprocessed <- readRDS("df.combined.Rds")
df.combined.preprocessed <- readRDS(params$data)
# Data with duplicates removed and years limited
df0 <- df.combined.preprocessed %>% filter(!remove) %>%
	   			    filter(publication_year >= min.year &
					   publication_year <= max.year)

if (!params$catalogue == "all") {
  df0 <- filter(df0, catalog == params$catalogue)
}
```

---

## From particular data to global perspective?

**Idea**: Study early modern knowledge production, all of it (`r min.year`-`r max.year`)  
**How?** Combine library catalogues (Fennica, Kungliga, ESTC, CERL HPB database)

**Important**: cleaning up the data (90% of the effort!) and considering how particular data collections (Finnish National Bibliography, English Short-Title Catalogue etc.) have been formed.

**How to fail**: take worldcat or similar combination of datasets and analyse it as a whole without cleaning it up first.

**Overcoming a national gaze** by focus on knowledge production through more complex cross-national network of cities

---

### Data: bibliographic catalogues 1640-1828

**Fennica** Finnish National bibliography

- `r nrow(subset(df0, catalog == "Fennica" & publication_year >= 1640 & publication_year < 1828))` documents 1640-1828

**Kungliga**  Swedish National bibliography

- `r nrow(subset(df0, catalog == "Kungliga" & publication_year >= 1640 & publication_year < 1828))` documents 1640-1828

**Why this project?**

- Visible roots of the diversity of "Finnishness" reflected in knowledge production, 1640-1828

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=3, echo=FALSE}
library(dplyr)
library(ggplot2)
theme_set(theme_bw(20))
df <- df0 %>% filter(publication_year >=1640 & publication_year <=1828) %>% group_by(catalog, publication_decade) %>% tally()
p <- ggplot(df, aes(x = publication_decade, y = n, color = catalog)) + 
       geom_point() + 
       geom_line() + 
       ylab("Documents (n)") + xlab("Publication year") +
       scale_color_manual(values = c("blue", "darkgreen")) + 
       guides(color = guide_legend(reverse = TRUE, title = "")) #+ 
print(p)
```

---

## Methods: Open data analytical ecosystem

**Customized** our dedicated R packages for bibliographic analysis combine flexibility with automation

**Automated** scales well with new data and unforeseen changes; a million documents cleaned up in 30 minutes

**Open source** fully transparent and reproducible data analytics: https://github.com/rOpenGov/fennica

**Research potential** greatly underestimated; world is full of metadata collections

![Data Science Workflow](data-science.png)  
\tiny{(Figure by Hadley Wickham)}

---

## Automated algorithms for data clean-up

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(bibliographica, quiet = TRUE)
```

**Clean up messy entries**

```{r, message=FALSE, warning=FALSE,echo=TRUE}
polish_physical_extent("iii-xxiv, 118, [2] p.")$pagecount
```

**Enrich data** (geocoordinates, gender, ..)

```{r, message=FALSE, warning=FALSE, echo=TRUE}
get_country("Porvoo")
```

- Open source tools: parse, clean up, interpret, augment, enrich,
  validate, summarise, analyze, visualize, report..

---

\begin{center}
\Large{1. General trends and political events}
\end{center}

---

\begin{center}
\tiny{\textcolor{gray}{General trends \& political events}}
\end{center}

### Turku as a Swedish University Town among Uppsala, Tartu and Stockholm (+ Lund & Greifswald)

```{r map1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=20, fig.height=4, eval=TRUE}
library(dplyr)
library(ggmap)
library(ggplot2)
library(gisfin)
source("funcs.R")
theme_set(get_theme_map())
i <- 1650
mapdata1 <- filter(df0,
              publication_decade == i &
              (catalog == "Kungliga" & 
	            publication_place %in% c("Stockholm", "Uppsala",
	                         "Greifswald", "Lund", "Tartu"))|
	  (catalog == "Fennica" &
	      publication_place == "Turku")) %>%
      group_by(publication_decade, publication_place,
	       latitude, longitude) %>%
      summarise(n = n())

@@

---

\begin{center}
\tiny{\textcolor{gray}{General trends \& political events}}
\end{center}

### Publishing activity over time

- 1700-1721 Great Northern War & Fennica drop during "the Great Wrath"
- 1760-1770 Pamphleteering plays a crucial role during abolished censorship
- 1803-1815 Napoleonic wars

```{r publishingovertime, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=4, out.width="300px"}

library(magrittr)
library(ggplot2)
library(dplyr)
df2 <- df0 %>% group_by(publication_decade, catalog) %>%
               summarise(n = n())
df2$highlight <- rep(FALSE, nrow(df2))
df2$highlight[df2$catalog == "Kungliga" &
	      df2$publication_decade %in% c(1770, 1810)] <- TRUE
df2$highlight[df2$catalog == "Fennica" &
	      df2$publication_decade %in% c(1710, 1810)] <- TRUE
v <- seq(1500, 1800, 100)
theme_set(theme_bw(20))
p1 <- ggplot(df2, aes(x = publication_decade, y = n)) +
     geom_line(aes(color = catalog), size = 1) +
     geom_point(aes(shape = catalog, color = catalog, size = highlight)) +
     ylab("Title count (n)") + xlab("Publication year") +
     #theme(legend.position = "none") +
     scale_x_continuous(breaks = v, labels = v) +
     scale_color_manual(values = c("blue", "darkgreen")) + 
  	 scale_size_discrete(range = c(2, 8)) +
     guides(color = guide_legend(reverse = TRUE), size = "none", shape = "none")

# Add indicator of the publication period
df <- df0 %>% filter(catalog == "Kungliga")
df$period <- cut(df$publication_year,
           breaks = c(1757, 1766, 1775, 1783),
		 include.lowest = TRUE,
		 right = FALSE
		 )

# Remove entries outside the investigated publication periods
df <- df %>% filter(!is.na(period)) %>% group_by(period, paper)
df <- df %>% summarise(n = n(),
		  paper2 = sum(paper, na.rm = TRUE)) %>%
             mutate(paper.per.title = paper2/n)		 
df$paper <- df$paper2
df$paper2 <- NULL

p2 <- ggplot(df, aes(x = period, y = paper.per.title)) + 
       geom_bar(stat = "identity") +
       scale_x_discrete(labels = c("1757-1765", "1766-1774", "1775-1783")) +
       ylab("Paper per title") + xlab("")

library(gridExtra)
grid.arrange(p1, p2, nrow = 1)
```

---

\begin{center}
\tiny{\textcolor{gray}{General trends \& political events}}
\end{center}

### Swedish knowledge production & cycles of representation

An example how historical events, namely **parliamentary assemblies** in the Swedish case, have an effect on knowledge production.

```{r riksdar5, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}

minyear <- 1700
  maxyear <- 1800
  catalog <- "Kungliga"

library(magrittr)
library(ggplot2)
library(dplyr)
library(ggplot2)
library(reshape2)

#for (catal in as.character(unique(df0$catalog))) {

for (catal in "Kungliga") {

  # Use timeinterval year intervals
  dfsel = subset(df0, catalog == catal)

  df <- df0
  df = subset(df, catalog == catal)
  df <- subset(df, publication_year >= minyear & publication_year <= maxyear)

  df <- subset(dfsel, publication_year >= 1700 & publication_year < 1800 )

  timeinterval <- 1
  df$timeunit <- round(df$publication_year/timeinterval)*timeinterval

  df$unity = rep(1, nrow(df))
  publications <- tapply(df$unity, list(df$timeunit), sum)

  publications[is.na(publications)] <- 0 # Set NAs to 0
  publications <- publications/timeinterval # Instead of decadal sum, use average annual output 
  dfm <- melt(publications) 
  names(dfm) <- c("Time", "Documents")
  dfm <- transform(dfm, date = as.numeric(as.character(Time)))
  ymin = min(dfm$Documents)
  ymax = max(dfm$Documents)

  rect_left <- c(min(na.omit(dfm$date)),
               1719-.5, 1719+.5, # Stockholm 20 januari 1719 1 juni 1719
               1734-.5, 1734+.5, # Stockholm 14 maj 1734 14 december 1734
               1765-.5, 1766+.5, # Stockholm 21 februari 1765 21 oktober 1766
               1769-.5, 1770+.5, # Norrköping & Stockholm 22 april 1769 5 februari 1770
               1771-.5, 1772+.5, # Stockholm 19 juni 1771 12 september 1772
               1792-.5, 1792+.5, # Gävle 26 januari 1792 24 februari 1792         
               max(na.omit(dfm$date)))
  rectangles <- data.frame(
    xmin = rect_left[-length(rect_left)],
    xmax = rect_left[-1],
    ymin = ymin,
    ymax = ymax
    )
  rectangles$shade <- rep(c("Background", "Highlight"), length = nrow(rectangles))

 riksplace = c("Stockholm", "Stockholm", "Stockholm", "Norrköping", "Stockholm", "Gävle")
 cols = c("white", "Purple", "yellow", "orange")
 rectangles$shade[rectangles$shade == "Highlight"] = riksplace
 rectangles$shade = factor(rectangles$shade)

  # Draw Figure
  theme_set(theme_bw(20))
  p <- ggplot()
  p <- p + geom_rect(data = rectangles, 
       aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax, fill=shade), alpha=0.8) + 
         scale_fill_manual(values = cols) # + guides(fill = "none") 
  p <- p + geom_line(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + geom_point(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + scale_x_continuous(breaks=seq(min(dfm$date), max(dfm$date), 20))
  p <- p + ggtitle("Publishing activity")
  p <- p + ylab("Documents / Year")
  p <- p + xlab("Year")
  p <- p + ggtitle(paste(catal,"Riksplaces"))
  print(p)
}

@@

---

\begin{center}
\Large{2. Publishers in Sweden and Turku}
\end{center}

---

\begin{center}
\tiny{\textcolor{gray}{Publishers in Sweden and Turku}}
\end{center}

### Turku in Sweden: proportions for title count and paper in Swedish University towns

```{r comparisons, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=18, fig.height=10}
pics <- list()
for (catalogue in na.omit(unique(df0$catalog))) {

df <- df0 %>% filter(catalog == catalogue)%>%
        filter(publication_year >= 1640 & publication_year <= 1828)
places <- c("Stockholm", "Uppsala", "Lund", "Greifswald", "Tartu", "Turku")
df$publication_place[!df$publication_place %in% places] <- "Other"
df <- subset(df, !is.na(df$publication_place))
#df$publication_place <- factor(df$publication_place, levels = c(places, "Other"))
library(reshape2)
library(magrittr)
library(dplyr)
library(tidyr)

df <- df %>% group_by(publication_decade, publication_place) %>%
             summarise(n = n(),
                paper = sum(paper, na.rm = TRUE),
		       publishers = length(na.omit(unique(publisher)))
		       )

# Calculate percentages
for (varname in c("n", "paper", "publishers")) {

  df$varname <- df[[varname]]
  dff <- df %>% select(publication_decade, publication_place, varname)
  dff <- spread(dff, publication_place, varname, fill = 0)
  dff[, -1] <- 100 * t(apply(dff[, -1], 1, function (x) {x/sum(x)}))
  dff <- melt(dff, "publication_decade");
  colnames(dff) <- c("publication_decade", "publication_place", "f")

  theme_set(theme_bw(20))
  p <- ggplot(dff, aes(x = publication_decade, y = f)) +
     geom_bar(position = "stack", stat = "identity",
       aes(fill = publication_place)) + 
     xlab("Publication year") +
     scale_fill_brewer(palette="Spectral")
     if (varname == "n") {
       p <- p + ggtitle(paste("Title count (", catalogue, ")", sep = "")) + 
                ylab("Title count frequency (%)") 
     }
     if (varname == "paper") {
       p <- p + ggtitle(paste("Paper (", catalogue, ")", sep = "")) +
                ylab("Paper (%)")        
     }
     if (varname == "publishers") {
       p <- p + ggtitle(paste("Publishers (", catalogue, ")", sep = "")) +
                       ylab("Publishers (%)")
     }     
  p <- p + guides(fill = guide_legend(reverse = TRUE, title = ""))

  pics[[paste(catalogue, varname, sep = "-")]] <- p

}
}
library(gridExtra)
grid.arrange(pics[[1]], pics[[2]], pics[[4]], pics[[5]], nrow = 2)
rm(pics)
```

---

\begin{center}
\tiny{\textcolor{gray}{Publishers in Sweden and Turku}}
\end{center}

### Publishers in Swedish University towns (and Finland)

```{r publishers2-finland, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
df <- df0 %>%
        filter(catalog == "Fennica") %>%
	filter(country == "Finland" | publication_place == "Vyborg") %>%
        filter(publication_year >= 1640 & publication_year <= 1828)

# Top publication places
library(bibliographica)

top <- c("Turku", "Vaasa", "Vyborg")
df <- df %>%	
        filter(publication_place %in% top) %>%	
        select(publication_decade, publication_place, publisher)

npub <- unique(df) %>% group_by(publication_decade, publication_place) %>% tally()
npub$publication_place <- factor(npub$publication_place, levels = top)
npub.fennica <- npub

df <- df0
catalogue <- "Kungliga"
df <- df %>%
        filter(catalog == catalogue) %>%
        filter(publication_year >= 1640 & publication_year <= 1828)

# Selected publication places
top <- c("Stockholm", "Lund", "Uppsala", "Greifswald", "Tartu")
df <- df %>%	
        filter(publication_place %in% top) %>%	
        select(publication_decade, publication_place, publisher)

npub <- unique(df) %>% group_by(publication_decade, publication_place) %>% tally()
npub$publication_place <- factor(npub$publication_place, levels = top)
npub.kungliga <- npub

npub <- bind_rows(npub.fennica, npub.kungliga)

# Order by total publishing activity
npub$publication_place <- factor(npub$publication_place, levels = (npub %>% group_by(publication_place) %>% summarise(total = sum(n)) %>% arrange(desc(total)))$publication_place)
npub <- npub %>% arrange(publication_place)

theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack",
       		    aes(fill = publication_place)) + 
       xlab("Publication year") +
       ylab("Unique publishers (n)") +
      scale_fill_brewer(palette="Spectral") +       
       guides(fill = guide_legend(reverse = TRUE, title = "")) #+ 
       #ggtitle(paste("Unique publishers in selected publication places (", catalogue, ")"))

print(p)
```

---

\begin{center}
\tiny{\textcolor{gray}{Publishers in Sweden and Turku}}
\end{center}

### Top publishers in Turku/Fennica

```{r publishers4-fennica, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=6, fig.width=12}
library(dplyr)
place <- "Turku"
catalogue <- "Fennica"

df <- df0 %>% filter(catalog == catalogue & publication_place == place)

df <- df %>%	
  filter(publication_year >= 1640 & publication_year <= 1828)%>%
        select(publication_decade, publisher, paper)

# Group small publishers
# Top publishers by title count
ntop <- 10
top <- names(top(df, "publisher", ntop))
df$publisher[!df$publisher %in% top] <- "Other"
df$publisher <- factor(df$publisher, c(top, "Other"))

# Title count per decade & publisher	
npub <- df %>% group_by(publication_decade, publisher) %>% tally() %>% arrange(n)
#npub$publisher <- factor(npub$publisher, levels = (npub %>% group_by(publisher) %>% summarise(total = sum(n)) %>% arrange(total))$publisher)
# Sort by biggest publication decade
#ord <- (npub %>% group_by(publisher) %>% summarise(decade = publication_decade[[which.max(n)]]) %>% arrange(decade))$publisher
ord <- as.character((npub %>% group_by(publisher) %>% 
                        summarise(decade = median(publication_decade)) %>% 
                        arrange(decade))$publisher)
npub$publisher <- factor(npub$publisher, levels = c("Other", setdiff(ord, "Other")))

# TITLE COUNT
theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = publisher)) +
      geom_point(aes(size = n)) + 
      #geom_line(aes(color = publisher), size = 2) +
      #geom_smooth(aes(fill = publisher)) +
      # geom_bar(stat = "identity", position = "stack", aes(fill = publisher)) + 
       xlab("Year") +
       ylab("") +
       guides(size = guide_legend(title = "Title count (n)"))
       #ggtitle(paste("Title count per publisher (", place, "/", catalogue, ")", sep = ""))
print(p)
```

---

\begin{center}
\Large{3. Languages and religious texts in Finland}
\end{center}

---

\begin{center}
\tiny{\textcolor{gray}{Languages \& religious texts in Finland}}
\end{center}

### Different languages in Turku, Stockholm, Uppsala and Lund

```{r language-perc2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=20, fig.height=12}
pics <- list()
theme_set(theme_bw(20))
myplaces <- c("Turku", "Greifswald", "Stockholm", "Uppsala", "Lund")

for (catalogue in c("Fennica", "Kungliga")) {

  df <- df0 %>% filter(catalog == catalogue) %>%
        filter(publication_year >= 1600 & publication_year <= 1828)

  langs <- c("Finnish", "Swedish", "Latin", "German",
           "Russian", "French", "Hebrew", "Other")

  lang <- paste("language.", langs, sep = "")
  otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
  otherlang <- otherlang[otherlang!="language.Multiple languages"]

  df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0

  dfl <- NULL
  for (lan in lang) {

    # Classify a document to the specifed language
    # If document is assigned with languages, each case is considered
    # so one doc may have multiple entries corresponding to different languages
    # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
    # Combine data frames for specified languages
    dflsub <- filter(df, df[[lan]])
    if (nrow(dflsub) > 0) {
      dflsub$language <- gsub("language.", "", lan)
      dfl <- bind_rows(dfl, dflsub)
    }
  }

  df <- dfl %>% group_by(publication_decade, publication_place, language) %>%
     	     summarise(n = n())

  # Calculate percentages
  dff <- spread(df, language, n, fill = 0)
  dff[, -c(1,2)] <- 100 * t(apply(dff[, -c(1,2)], 1, function (x) {x/sum(x)}))

  dff <- melt(dff, id = c("publication_place", "publication_decade"))
  colnames(dff) <- c("publication_place", "publication_decade", "language", "f")

  for (myplace in myplaces) {
    p <- ggplot(subset(dff, publication_place == myplace), aes(x = publication_decade, y = f)) +
      geom_bar(position = "stack", stat = "identity", aes(fill = language)) + 
      xlab("Publication year") +
      ylab("Title count frequency (%)") +
      #ggtitle(paste("Languages (", myplace, "/", catalogue, ")", sep = "")) +
      ggtitle(paste(myplace, "/", catalogue)) +
      scale_fill_brewer(palette="Spectral") + 
      guides(fill = guide_legend(reverse = TRUE, title = ""))
    pics[[paste(catalogue, myplace, sep = "-")]] <- p
  }
}

grid.arrange(pics[[1]],
             pics[[8]], pics[[9]], pics[[10]], 
             nrow = 2)

rm(pics)
```

---

\begin{center}
\tiny{\textcolor{gray}{Languages \& religious texts in Finland}}
\end{center}

### Topic richness per language (publications in Fennica)

```{r topics232, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, fig.show="hold"}
catalogue <- "Fennica"
df <- filter(df0, catalog == "Fennica") %>%
  filter(publication_year >= 1640 & publication_year <= 1828) %>%
  select(publication_year, language, subject_topic)
df <- unique(df)
df <- df %>% group_by(publication_year, language) %>%
             summarise(n = length(unique(unlist(strsplit(subject_topic, ";")))))

langs2 <- c("Finnish", "Swedish", "Latin", "Other")
df$language[!df$language %in% langs2] <- "Other"
df$language <- factor(df$language, levels = langs2)

p <- ggplot(df, aes(x = publication_year, y = n, color = language)) +
       geom_line() + geom_point() + 
       geom_smooth(aes(fill = language, color = language)) + 
       ylab("Documents (n)") + xlab("Publication year")

print(p)
```

---

\begin{center}
\tiny{\textcolor{gray}{Languages \& religious texts in Finland}}
\end{center}

### Devotional literature (catechisms, hymns, prayers, etc.)

Form of literature denoting the advancement of reading in Finland in the 19th century. It has been unclear when devotional literature  shows up as a relevant category.

```{r topics12122, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, fig.show="hold"}
sel <- c("virret","arkkiveisut","hartauskirjat","katekismukset","rukouspäivät","saarnat","aapiset","rukoukset","rukous","hengelliset laulut","hartauspuheet","virsikirjat")
catalogue <- "Fennica"

  # Selected catalogue with selected years
  df <- filter(df0, catalog == catalogue &
     		    publication_year >= 1640 &
		    publication_year <= 1828)

  # Selected topics
  df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
  df <- df %>% filter(catalog == catalogue & hit)

  lang <- paste("language.", langs, sep = "")
  otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
  df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
  dfl <- NULL
  for (lan in lang) {
    # Classify a document to the specifed language
    # If document is assigned with languages, each case is considered
    # so one doc may have multiple entries corresponding to different languages
    # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
    # Combine data frames for specified languages
    dflsub <- filter(df, df[[lan]])
    if (nrow(dflsub) > 0) {
      dflsub$language <- gsub("language.", "", lan)
      dfl <- bind_rows(dfl, dflsub)
    }
  }
  df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

  # PAPER CONSUMPTION
  theme_set(theme_bw(20))
  p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       #geom_smooth(aes(fill = language, color = language)) + 
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Devotional Literature (", catalogue, ")", sep = ""))
       #ggtitle(paste("Devotional literature" (", catalogue, ")", sep = "")) 
  print(p)
```

---

\begin{center}
\tiny{\textcolor{gray}{Languages \& religious texts in Finland}}
\end{center}

### Temporary sermons in Latin and Swedish

- Important role in early modern knowledge production. Language of temporary sermons in contrast to devotional literature has not been analysed in Finland (and Sweden).

```{r topics882, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
  sel <- c("hautajaiset", "häät", "juhlamenot")
  catalogue <- "Fennica"
  df <- df0
  df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)

  # Selected catalogue with selected topics
  df <- df %>% filter(catalog == catalogue & hit)
  lang <- paste("language.", langs, sep = "")
  otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
  otherlang <- otherlang[otherlang!="language.Multiple languages"]
  df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0

  dfl <- NULL
  for (lan in lang) {
    # Classify a document to the specifed language
    # If document is assigned with languages, each case is considered
    # so one doc may have multiple entries corresponding to different languages
    # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
    # Combine data frames for specified languages
    dflsub <- filter(df, df[[lan]])
    if (nrow(dflsub) > 0) {
      dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }
}
dfl$language <- factor(dfl$language, levels = langs)

  df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

  # PAPER CONSUMPTION
  p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Temporary Sermons (", catalogue, ")", sep = ""))
  print(p)
```

---

\begin{center}
\Large{4. University curriculum and Enlightenment in Finland}
\end{center}

---

\begin{center}
\tiny{\textcolor{gray}{University curriculum \& Enlightenment in Finland}}
\end{center}

### Dissertation languages in Turku 1640-1828

```{r disslang, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
catalogue <- "Fennica"
langs <- c("Finnish", "Swedish", "Latin", "German",
           "Russian", "French", "Hebrew", "Other")

df <- df0 %>%
        filter(catalog == catalogue) %>%
        filter(publication_year >= 1640 & publication_year <= 1828) %>%
	filter(dissertation)

lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
otherlang <- otherlang[otherlang!="language.Multiple languages"]

df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0

dfl <- NULL
for (lan in lang) {

  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  if (nrow(dflsub)>0) {
    dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }
}

df <- dfl %>% group_by(publication_decade, language) %>%
     	     summarise(n = n())

# Calculate percentages
dff <- spread(df, language, n, fill = 0)
dff[, -1] <- 100 * t(apply(dff[, -1], 1, function (x) {x/sum(x)}))
dff <- melt(dff, "publication_decade");
colnames(dff) <- c("publication_decade", "language", "f")

# Order by language counts
dff$language <- factor(dff$language,
   levels = (df %>% group_by(language) %>%
                    summarise(total = sum(n)) %>%
		    arrange(total))$language);
dff <- dff %>% arrange(desc(language))

theme_set(theme_bw(20))
p <- ggplot(dff, aes(x = publication_decade, y = f)) +
     geom_bar(position = "stack", stat = "identity", aes(fill = language)) + 
     xlab("Publication year") +
     ylab("Title count frequency (%)") +
     ggtitle(paste("Dissertation languages (", catalogue, ")", sep = "")) +
     scale_fill_brewer(palette="Spectral")
print(p)
```

---

\begin{center}
\tiny{\textcolor{gray}{University curriculum \& Enlightenment in Finland}}
\end{center}

### Printing by topics in different languages in Turku

```{r topics-103B, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=8}
sel = c("filosofia","luonnonfilosofia","metafysiikka","logiikka")
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
                 paper = sum(paper, na.rm = TRUE))

# PAPER CONSUMPTION
p1 <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_bar(aes(fill = language), position = "stack", stat = "identity") +                     
       xlab("Publication year") +
       ylab("Paper") +
       guides(fill = guide_legend(reverse = TRUE, title = "")) +
       ggtitle("Philosophy")

sel = c("oppihistoria","antiikki","historia")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

# PAPER CONSUMPTION
p2 <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_bar(aes(fill = language), position = "stack", stat = "identity") +             
       xlab("Publication year") +
       ylab("Paper") +
       guides(fill = guide_legend(reverse = TRUE, title = "")) +
       ggtitle("History")

sel =   c("psykologia","yhteiskuntafilosofia","valtiofilosofia","talous","raha","moraali","velvollisuudet","kasvatus","maanviljely","maatalous","kalastus")

>>>>

---
title: "Untitled"
output: 
  flexdashboard::flex_dashboard:
    logo: ropengov.png
    favicon: HY-CC.png
    theme: default
    orientation: columns
    vertical_layout: fill
    navbar:
      - { title: "About", href: "https://example.com/about", align: left }
    social: [ "menu" ]
    source_code: http://github.com/antagomir
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
```

```{r init, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE}
library(ggmap)
library(gisfin)
library(stringr)
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(bibliographica)
library(fennica)
library(sorvi)
library(devtools)
# install_github("ropengov/gisfin")
# devtools::install_github('cttobin/ggthemr')
#library(ggthemr)
knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(fig.path = "figure_slides/", dev="CairoPNG")
knitr::opts_chunk$set(fig.path = "figure_flex/")

@@

# Params
min.year <- 1470
max.year <- 1828

# Full combined catalogue (Fen + Kun) with marked duplicates
df.combined.preprocessed <- readRDS("df.combined.Rds")
# Data with duplicates removed and years limited
df0 <- df.combined.preprocessed %>% filter(!remove) %>%
	   			    filter(publication_year >= min.year &
					   publication_year <= max.year)
```

Page 1
==============================================================================

Column {data-width=650}
-----------------------------------------------------------------------

Flexdashboard instructions:
  * [Usage](http://rmarkdown.rstudio.com/flexdashboard/using.html)
  * [htmlwidget gallery](http://gallery.htmlwidgets.org/)
  * [flexdashboard examples](http://rmarkdown.rstudio.com/flexdashboard/examples.html)
  * [plotly](https://plot.ly/r/)

### Linked Value Box

```{r}
valueBox(42, icon = "fa-pencil", href="#page-2")
```

### Table

```{r}
library(shiny)
#renderTable({head(mtcars)})
#renderTable({head(mtcars, n = input$rows)})

library(DT)
DT::datatable(mtcars, options = list(bPaginate = FALSE))
#DT::datatable(mtcars, options = list(pageLength = 25))
```

<!--Column {data-width=350}-->
Column {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Chart B

```{r map1, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggmap)
library(ggplot2)
library(gisfin)
source("funcs.R")
theme_set(get_theme_map())
i <- 1650
mapdata1 <- filter(df0,
              publication_decade == i &
              (catalog == "Kungliga" & 
	            publication_place %in% c("Stockholm", "Uppsala",
	                         "Greifswald", "Lund", "Tartu"))|
	  (catalog == "Fennica" &
	      publication_place == "Turku")) %>%
      group_by(publication_decade, publication_place,
	       latitude, longitude) %>%
      summarise(n = n())

@@

### Chart C

```{r}
library(dygraphs)
dygraph(ldeaths)
```

Page 2
=====================================================================

Column {data-width=650}
-----------------------------------------------------------------------

### Chart A

```{r}

@@

Column {data-width=350}
-----------------------------------------------------------------------

### Chart B

```{r}

### Chart C

```{r}
